spinifex: An R Package for Creating a Manual Tour of Low-dimensional Projections of Multivariate Data

### 2020-09-12 paper/spyrison-cook.Rmd
final fixes for the published version

knitr::opts_chunk$set(
  fig.align  = "center",
  echo       = TRUE,
  collapse   = TRUE,
  message    = FALSE,
  warning    = FALSE,
  error      = FALSE,
  cache      = FALSE,
  cache.lazy = FALSE
)

@@

```{r step2, echo=F, out.width='70%',                                                fig.cap="Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data. The blue circle and variable map lies on the projection plane. The red circle, orthogonal to the projection plane, illustrates the manipulation space and how the manip var can be controlled and how this affects the variable contribution back onto the projection plane. The other variables are omitted from the manipulation dimension for simplicity. Picturing the other variables in that dimension reveals the intuition that rotating one variable performs a constrained rotation on the others. This is illustrated with the \\code{view\\_manip\\_space()} function."}
view_manip_space2(basis = f_bas, manip_var = f_mvar, lab = colnames(f_dat))
```

@@

```{r step3, echo=F, warning=F, out.width='100%',                                     fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original. "}
f_angle <- .29
f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
if(dim(f_mtour)[3] != 15) message(
  paste0("!!! n_frames != 15 !!! Check angle step size. n_frames = ", dim(f_mtour)[3], ". angle = ", f_angle))
ggproto
z <- data.frame(0,0,0,0,0,0)

@@

Tours are typically viewed as an animation. The animation of this tour can be viewed online at \url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The page may take a moment to load. Animations can be produced using the function \code{play\_manual\_tour()}.

```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in',                 fig.cap="Radial manual tour manipulating aede2 of standardized flea data. The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value. This effects the separation between the purple and green clusters. This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection. An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
## Figure
array2ggfacets(tour_array = f_mtour, data = f_dat, m_var = f_mvar, class = f_clas)

@@

<!-- discussion of findings and which is more insightful -->
When PC4 is removed from the projection (Figure \ref{fig:JetClusterGood}) the difference between the two groups is removed, indicating that it is important for distinguishing experiments. However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}) does not affect the structure, indicating it is not important for distinguishing experiments. Animations for the remaining PCs can be viewed at the following links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/). It can be seen that only PC4 is important for viewing the difference in these two experiments.

```{r JetClusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in',        fig.cap="Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange). When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments. The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc4/."}
# Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
load("./data/JetCluster_sub.rda")
load("./data/JetCluster_basis.rda")

@@

```{r JetClusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',         fig.cap="Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).  When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments. The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\\_manualtour\\_pc3/."}
# Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
jet_mvar  <- 3
jet_mtour <- manual_tour(basis = jet_bas, manip_var = jet_mvar, angle = jet_ang)

@@

<!-- comparison of the DIS cluster -->
Both PC2 and PC6 contribute to the projection similarly. When PC6 is rotated into the projection, variation in the DIS HERA1+2 is greatly reduced. When PC2 is removed from the projection, dimuon SIDIS becomes more clearly distinct. Even though both variables contribute similarly to the original projection, their contributions have quite different effects on the structure of each cluster, and the distinction between clusters. Animations of all of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).

```{r DISclusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in',        fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2. As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6. The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc6/."}

@@

```{r DISclusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in',         fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important. The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\\_manualtour\\_pc2/."}
##DIS cluster, fig 8 of cook_dynamical_2018
DIS_mvar  <- 2
DIS_mtour <- manual_tour(basis = DIS_bas, manip_var = DIS_mvar, angle = DIS_ang)

@@

This article was created in R, using \CRANpkg{knitr} [@xie_knitr_2020] and \CRANpkg{rmarkdown} [@allaire_rmarkdown_2020], with code generating the examples inline. The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/). The source code for the \pkg{spinifex} package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

>>>>

knitr::opts_chunk$set(
  fig.align  = "center",
  echo       = TRUE,
  collapse   = TRUE,
  message    = FALSE,
  warning    = FALSE,
  error      = FALSE,
  cache      = FALSE,
  cache.lazy = FALSE,
  fig.retina = 3
)

@@

```{r step2, echo=F, out.width='50%', fig.width=3, fig.height=3,  fig.cap="Illustration of a 3D manip space, this space is rotated effectively changing the contribution of the manip variable, aede2 in the example data. The blue circle and variable map lies on the projection plane. The red circle, orthogonal to the projection plane, illustrates the manipulation space and how the manip var can be controlled and how this affects the variable contribution back onto the projection plane. The other variables are omitted from the manipulation dimension for simplicity. Picturing the other variables in that dimension reveals the intuition that rotating one variable performs a constrained rotation on the others. This is illustrated with the \\code{view\\_manip\\_space()} function."}
view_manip_space2(basis = f_bas, manip_var = f_mvar, lab = colnames(f_dat))
```

@@

```{r step3, echo=F, warning=F, out.width='100%', fig.width=8, fig.height=2, fig.cap="Snapshots of a radial manual tour manipulating aede2: (1) original projection, (2) full contribution, (3) zero contribution, (4) back to original. "}
f_angle <- .29
f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
if(dim(f_mtour)[3] != 15) message(
  paste0("!!! n_frames != 15 !!! Check angle step size. n_frames = ", dim(f_mtour)[3], ". angle = ", f_angle))
ggproto
z <- data.frame(0,0,0,0,0,0)

@@

Tours are typically viewed as an animation. The animation of this tour can be viewed online at \url{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/flea_radialtour_mvar5.gif}. The page may take a moment to load. Animations can be produced using the function \code{play\_manual\_tour()}.

```{r step4, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Radial manual tour manipulating aede2 of standardized flea data. The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value. This effects the separation between the purple and green clusters. This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection."}
## Figure
array2ggfacets(tour_array = f_mtour, data = f_dat, m_var = f_mvar, class = f_clas)

@@

<!-- discussion of findings and which is more insightful -->
When PC4 is removed from the projection (Figure \ref{fig:JetClusterGood}) the difference between the two groups is removed, indicating that it is important for distinguishing experiments. However, removing PC3 from the projection (Figure \ref{fig:JetClusterBad}) does not affect the structure, indicating it is not important for distinguishing experiments. Animations for the remaining PCs can be viewed at the following links: [PC1](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc1.gif), [PC2](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc2.gif), [PC3](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc3.gif), and [PC4](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc4.gif). It can be seen that only PC4 is important for viewing the difference in these two experiments.

```{r JetClusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange). When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments."}
# Jet cluster, Fig 7 of cook_dynamical_2018, subset to ATLAS7old and ATLAS7new
load("./data/JetCluster_sub.rda")
load("./data/JetCluster_basis.rda")

@@

```{r JetClusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).  When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments."}
# Jet cluster, Fig 7 of cook_dynalical_2018, subset to ATLAS7old and ATLAS7new
jet_mvar  <- 3
jet_mtour <- manual_tour(basis = jet_bas, manip_var = jet_mvar, angle = jet_ang)

@@

<!-- comparison of the DIS cluster -->
Both PC2 and PC6 contribute to the projection similarly. When PC6 is rotated into the projection, variation in the DIS HERA1+2 is greatly reduced. When PC2 is removed from the projection, dimuon SIDIS becomes more clearly distinct. Even though both variables contribute similarly to the original projection, their contributions have quite different effects on the structure of each cluster, and the distinction between clusters. Animations of all of the principal components can be viewed from the links: [PC1](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc1.gif), [PC2](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc2.gif), [PC3](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc3.gif), [PC4](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc4.gif), [PC5](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc5.gif), and [PC6](https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc6.gif).

```{r DISclusterGood, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2. As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6."}

@@

```{r DISclusterBad, echo=F, warning=F, out.height='7in', out.width='5.83in', fig.cap="Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important."}
##DIS cluster, fig 8 of cook_dynamical_2018
DIS_mvar  <- 2
DIS_mtour <- manual_tour(basis = DIS_bas, manip_var = DIS_mvar, angle = DIS_ang)

@@

This article was created in R, using \CRANpkg{knitr} [@xie_knitr_2020] and \CRANpkg{rmarkdown} [@allaire_rmarkdown_2020], with code generating the examples inline. The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/). The animated gifs can also be viewed at this site, and also in the supplementary material for this paper. The source code for the \pkg{spinifex} package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

### 2020-09-12 paper/spyrison-cook.tex
final fixes for the published version

% !TeX root = RJwrapper.tex
\title{spinifex: An R Package for Creating a Manual Tour of Low-dimensional
Projections of Multivariate Data}
\author{by Nicholas Spyrison, Dianne Cook}

@@

{\centering \includegraphics[width=0.7\linewidth]{spyrison-cook_files/figure-latex/step2-1}

@@

@@

Tours are typically viewed as an animation. The animation of this tour
can be viewed online at
\url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
page may take a moment to load. Animations can be produced using the
function \code{play\_manual\_tour()}.

@@

\caption[Radial manual tour manipulating aede2 of standardized flea data]{Radial manual tour manipulating aede2 of standardized flea data. The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value. This effects the separation between the purple and green clusters. This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection. An animation can be viewed at https://nspyrison.netlify.com/thesis/flea\_manualtour\_mvar5/.}\label{fig:step4}
\end{figure}
\end{Schunk}

@@

\caption{\label{tab:functionsTable}Summary of available functions.}
\centering
\begin{tabular}[t]{lll}
\toprule
Type & Function & Description\\
\midrule
construction & create\_manip\_space & forms the 3D space of rotation\\
construction & rotate\_manip\_space & performs 3D rotation\\
construction & manual\_tour & generates sequence of 2D frames\\
 &  & \\
render & array2df & turn the tour path array into long form, for plotting\\
render & render\_ & render long form as a ggplot2 objection for animation\\
render & render\_plotly & render the animation as a plotly object (default)\\
render & render\_gganimate & render the animation as a gganimate object\\
 &  & \\
animation & play\_tour\_path & composite function animating the specified tour path\\
animation & play\_manual\_tour & composite function animating the specified manual tour\\
 &  & \\
specialty & oblique\_basis & table of the rotated basis and manip space\\
specialty & oblique\_frame & display the reference axes of a given basis\\
specialty & view\_manip\_space & illustrative display of any manip space\\
\bottomrule
\end{tabular}
\end{table}

@@

When PC4 is removed from the projection (Figure
\ref{fig:JetClusterGood}) the difference between the two groups is
removed, indicating that it is important for distinguishing experiments.
However, removing PC3 from the projection (Figure
\ref{fig:JetClusterBad}) does not affect the structure, indicating it is
not important for distinguishing experiments. Animations for the
remaining PCs can be viewed at the following links:
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2},
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3},
and
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.
It can be seen that only PC4 is important for viewing the difference in
these two experiments.

@@

\caption[Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange). When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments. The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc4/.}\label{fig:JetClusterGood}
\end{figure}
\end{Schunk}

@@

\caption[Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).  When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments. The animation can be viewed at https://nspyrison.netlify.com/thesis/jetcluster\_manualtour\_pc3/.}\label{fig:JetClusterBad}
\end{figure}
\end{Schunk}

@@

Both PC2 and PC6 contribute to the projection similarly. When PC6 is
rotated into the projection, variation in the DIS HERA1+2 is greatly
reduced. When PC2 is removed from the projection, dimuon SIDIS becomes
more clearly distinct. Even though both variables contribute similarly
to the original projection, their contributions have quite different
effects on the structure of each cluster, and the distinction between
clusters. Animations of all of the principal components can be viewed
from the links:
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5},
and
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.

@@

\caption[Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2. As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6. The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc6/.}\label{fig:DISclusterGood}
\end{figure}
\end{Schunk}

@@

\caption[Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important. The animation can be viewed at https://nspyrison.netlify.com/thesis/discluster\_manualtour\_pc2/.}\label{fig:DISclusterBad}
\end{figure}
\end{Schunk}

@@

This article was created in R, using \CRANpkg{knitr}
\citep{xie_knitr_2020} and \CRANpkg{rmarkdown}
\citep{allaire_rmarkdown_2020}, with code generating the examples
inline. The source files for this article be found at
\href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
The source code for the \pkg{spinifex} package can be found at
\href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.

@@

\address{%
Nicholas Spyrison\\
Monash University\\
Faculty of Information Technology\\
}
\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}

\address{%
Dianne Cook\\
Monash University\\
Department of Econometrics and Business Statistics\\
}
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}

>>>>

% !TeX root = RJwrapper.tex
\title{spinifex: An R Package for Creating a Manual Tour of Low-dimensional
Projections of Multivariate Data}
\author{by Nicholas Spyrison and Dianne Cook}

@@

{\centering \includegraphics[width=0.5\linewidth]{spyrison-cook_files/figure-latex/step2-1}

@@

\begin{Schunk}
\begin{Soutput}
#> function (`_class` = NULL, `_inherit` = NULL, ...) 
#> {
#>     e <- new.env(parent = emptyenv())
#>     members <- list(...)
#>     if (length(members) != sum(nzchar(names(members)))) {
#>         abort("All members of a ggproto object must be named.")
#>     }
#>     if (length(members) > 0) {
#>         list2env(members, envir = e)
#>     }
#>     `_inherit` <- substitute(`_inherit`)
#>     env <- parent.frame()
#>     find_super <- function() {
#>         eval(`_inherit`, env, NULL)
#>     }
#>     super <- find_super()
#>     if (!is.null(super)) {
#>         if (!is.ggproto(super)) {
#>             abort("`_inherit` must be a ggproto object.")
#>         }
#>         e$super <- find_super
#>         class(e) <- c(`_class`, class(super))
#>     }
#>     else {
#>         class(e) <- c(`_class`, "ggproto", "gg")
#>     }
#>     e
#> }
#> <bytecode: 0x7fcdaa8d99c8>
#> <environment: namespace:ggplot2>
\end{Soutput}
\begin{figure}

@@

Tours are typically viewed as an animation. The animation of this tour
can be viewed online at
\url{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/flea_radialtour_mvar5.gif}.
The page may take a moment to load. Animations can be produced using the
function \code{play\_manual\_tour()}.

@@

\caption[Radial manual tour manipulating aede2 of standardized flea data]{Radial manual tour manipulating aede2 of standardized flea data. The axis for aede2 increases in contribution to the projection, from its initial value to 1, decreasing to 0 and then returning to the initial value. This effects the separation between the purple and green clusters. This shows that aede2 is important for distinguishing the purple species, because the separation disappears when aede2 is not contributing to the projection.}\label{fig:step4}
\end{figure}
\end{Schunk}

@@

\caption{\label{tab:functionsTable}Summary of available functions.}
\centering
\begin{tabular}[t]{lll}
\toprule
Type & Function & Description\\
\midrule
construction & create\_manip\_space & forms the 3D space of rotation\\
construction & rotate\_manip\_space & performs 3D rotation\\
construction & manual\_tour & generates sequence of 2D frames\\
 &  & \\
render & array2df & turn the tour path array into long form, for plotting\\
render & render\_ & render long form as a ggplot2 objection for animation\\
render & render\_plotly & render the animation as a plotly object (default)\\
render & render\_gganimate & render the animation as a gganimate object\\
 &  & \\
animation & play\_tour\_path & composite function animating the specified tour path\\
animation & play\_manual\_tour & composite function animating the specified manual tour\\
 &  & \\
specialty & print\_manip\_space & table of the rotated basis and manip space\\
specialty & oblique\_frame & display the reference axes of a given basis\\
specialty & view\_manip\_space & illustrative display of any manip space\\
\bottomrule
\end{tabular}
\end{table}

@@

When PC4 is removed from the projection (Figure
\ref{fig:JetClusterGood}) the difference between the two groups is
removed, indicating that it is important for distinguishing experiments.
However, removing PC3 from the projection (Figure
\ref{fig:JetClusterBad}) does not affect the structure, indicating it is
not important for distinguishing experiments. Animations for the
remaining PCs can be viewed at the following links:
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc1.gif}{PC1},
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc2.gif}{PC2},
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc3.gif}{PC3},
and
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/jetcluster_manualtour_pc4.gif}{PC4}.
It can be seen that only PC4 is important for viewing the difference in
these two experiments.

@@

\caption[Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange). When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments.}\label{fig:JetClusterGood}
\end{figure}
\end{Schunk}

@@

\caption[Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type]{Snapshots of a radial manual tour of PC3 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange).  When the contribution from PC3 is changed there is little change to the structure of the two groups, suggesting that PC3 is not important for distinguishing the experiments.}\label{fig:JetClusterBad}
\end{figure}
\end{Schunk}

@@

Both PC2 and PC6 contribute to the projection similarly. When PC6 is
rotated into the projection, variation in the DIS HERA1+2 is greatly
reduced. When PC2 is removed from the projection, dimuon SIDIS becomes
more clearly distinct. Even though both variables contribute similarly
to the original projection, their contributions have quite different
effects on the structure of each cluster, and the distinction between
clusters. Animations of all of the principal components can be viewed
from the links:
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc1.gif}{PC1},
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc2.gif}{PC2},
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc3.gif}{PC3},
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc4.gif}{PC4},
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc5.gif}{PC5},
and
\href{https://github.com/nspyrison/spinifex_paper/blob/master/paper/gifs/discluster_manualtour_pc6.gif}{PC6}.

@@

\caption[Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC6 has on the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). DIS HERA1+2 is distributed in a cross-shaped plane, charm SIDIS occupies the center of this cross, and dimuon SIDIS is a linear cluster crossing DIS HERA1+2. As the contribution of PC6 is increased, DIS HERA1+2 becomes almost singular in one direction (frame 5), indicating that this experiment has very little variability in the direction of PC6.}\label{fig:DISclusterGood}
\end{figure}
\end{Schunk}

@@

\caption[Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type]{Snapshots of a radial manual tour exploring the sensitivity PC2 to the structure of the DIS cluster, with color indicating experiment type: DIS HERA1+2 (green), dimuon SIDIS (purple), and charm SIDIS (orange). As contribution from PC2 is decreased, dimuon SIDIS becomes more distinguishable from the other two clusters (frames 10-14), indicating that in its absence PC2 is important.}\label{fig:DISclusterBad}
\end{figure}
\end{Schunk}

@@

This article was created in R, using \CRANpkg{knitr}
\citep{xie_knitr_2020} and \CRANpkg{rmarkdown}
\citep{allaire_rmarkdown_2020}, with code generating the examples
inline. The source files for this article be found at
\href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
The animated gifs can also be viewed at this site, and also in the
supplementary material for this paper. The source code for the
\pkg{spinifex} package can be found at
\href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.

@@

\address{%
Nicholas Spyrison\\
Monash University\\%
Faculty of Information Technology\\
%
%
%
\\\href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}}
}

\address{%
Dianne Cook\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
%
%
\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
}
